Species persistence under climate and fishing

### 2014-02-03 Writing/MS.tex
reordered presentation of results in discussion

Knowing whether two disturbances are synergistic in their impacts for a given population is important for management. Climate change-driven range shifts and fishing form a potential, but largely unexamined, example of synergistic disturbances. We found a negative relationship between the critical harvesting rate and the rate of environmental shift.  That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct. The curved line separating parameters that will allow the population to persist from those that won't is an indication of an interaction between the stressors (Figure \ref{biomass}). We find that synergy between these two disturbances is greatest in the region of parameter space where the equilibrium biomass is smallest (Figure \ref{Synergy}). While synergistic effects were quite small, the synergy is additive, which means that harvesting levels or climate velocities that are sustainable individually together can drive a population to extinction. Further, the fact that synergy is highest in those populations whose persistence is most tenuous is worrisome if population persistence is a management objective. Although our model is quite general, this suggests that it is marginal populations that are most at risk from the synergistic effects of these two disturbances. 
% the below needs some help, Eleanor can you flesh out with a better discussion of how our results compare to Mora, Planque and Kirby?
Synergy between harvesting and climate changes, e.g. warming temperature or a more variable climate, has been identified in experiments \citep{Moraetal2007}, in the demography and size of specific populations \citep{Planque:2010uq}, and at the ecosystem level \citep{Kirby:2009fk}.

Our results also suggest that particular combinations of harvesting and rate of environmental shift will affect some species more than others. As shown in Figure \ref{baseline}, species with a shorter generation time and longer the average dispersal distance will better track a high rate of environmental shift relative to a species that has a long generation time and short dispersal distance. This is in agreement with empirical work which has found that fish which shifted in response to warming in North Sea had faster life histories than non shifting species (smaller body sizes, faster maturation, smaller sizes at maturity) \citep{Perryetal2005}.

@@

Previous modeling work examining examining how harvest and range shift intersect has lacked the generality to examine how these two disturbances are likely to jointly affect population persistence. Many modeling studies have focused on how fish populations are likely to change in the future under a variety of fishing and climate scenarios, but the degree of detail and case-specificity makes it difficult to draw general conclusions. Here we have built a general model to examine how climate and harvesting interact to affect species persistence by incorporating dispersal and reproduction.

The advantage of a simple model like ours is that it is general enough to be applied to a number of systems.  However, it  ignores many of the complexities present in marine fisheries. Allee effects, for example, have been suggested as an important factor in fish population dynamics. By leaving out Allee effects, populations which shrank to very low levels were still able for it to persist over time. However, we found that qualitatively similar results about the interaction between climate and harvesting would hold for a model with a recruitment function with an Allee effect; i.e. rather than extinction occurring when $n(t)=0$, it would be when $n(t)=\text{the Allee threshold}$. Thus the absolute values of the critical climate velocity and harvesting rates would change, but their relationship to each other and general shape would remain the same.

We also did not include age structure in our model. The effects of both harvesting and climate change may be different across different age classes which may cause long-term changes in abundance or increased variance due to cohort resonance \citep{Botsfordetal2011, Rouyeretal2012}. Both of these effects can be expected to reduce population viability in the face of both fishing and range shifts; including this level of complexity is left for future work. Additionally, we did not allow any mechanisms aside from larval dispersal by which the population could keep up with a shifting climate.  In addition to these species-specific extensions, this modeling framework could be extended to consider species interactions, especially predator-prey pairs.  By introducing a predatory species, we would be imposing yet another stressor on the focus species \citep{Lingetal2009, Gurevitchetal2000} and we are interested in measuring the interaction between the effects of this stressor and the two we consider here.

>>>>

Knowing whether two disturbances are synergistic in their impacts for a given population is important for management. The co-occurrence of climate change-driven range shifts and fishing mean that these is the potential for synergistic interactions, which have been largely unexamined.  Previous modeling work examining examining how harvest and range shift intersect has lacked the generality to examine how these two disturbances are likely to jointly affect population persistence. Many modeling studies have focused on how fish populations are likely to change in the future under a variety of fishing and climate scenarios, but the degree of detail and case-specificity makes it difficult to draw general conclusions. Here we have built a general model to examine how climate and harvesting interact to affect species persistence by incorporating dispersal and reproduction.

While we did find synergy between the stressors, the drop in biomass caused by both stressors was never very much higher than the null prediction, i.e. synergistic effects were quite small.  However, the fact that synergy is highest in those populations whose persistence is most tenuous is worrisome from a conservation perspective.  Additionally, the fact that we found additive synergy means that harvesting levels or rate of environment shift that are sustainable individually together can drive a population to extinction.  Synergy between harvesting and climate changes, e.g. warming temperature or a more variable climate, has been identified in experiments \citep{Moraetal2007}, in the demography and size of specific populations \citep{Planque:2010uq}, and at the ecosystem level \citep{Kirby:2009fk}.  Our results  agree with these empirical findings and it is therefore important to take the effects of both stressors into consideration when designing conservation and management strategies.
% the below needs some help, Eleanor can you flesh out with a better discussion of how our results compare to Mora, Planque and Kirby?

We also derived expressions for critical harvesting rates and critical rates of environmental shift for a separable Gaussian kernel and found these critical rates using numerical simulations of a Laplace kernel.  For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a population can adjust to harvest and climate change.  More interestingly, we found a negative relationship between the critical harvesting rate and the rate of environmental shift.  That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.  The curved line separating parameters that will allow the population to persist from those that won't is another indication of an interaction between the stressors.

Our results suggest that particular combinations of harvesting and rate of environmental shift will affect some species more than others. As shown in Figure \ref{baseline}, species with a shorter generation time and a longer average dispersal distance will better track a high rate of environmental shift relative to a species that has a long generation time and short dispersal distance. This is in agreement with empirical work which has found that fish which shifted in response to warming in North Sea had faster life histories than non shifting species (smaller body sizes, faster maturation, smaller sizes at maturity) \citep{Perryetal2005}.

@@
